---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# psHarmonize `r utils::packageVersion('psHarmonize')`

<!-- badges: start -->

<!-- badges: end -->

**psHarmonize** is an R package that facilitates harmonization of data from multiple different datasets.

The `harmonization()` function uses a **harmonization sheet** imported from .xlsx format as input into an R function. This sheet catalogs variable names, domains (e.g. clinical, behavioral, outcomes), provides R code instructions for mapping or conversion of data, specifies the variable name in the harmonized data set, and tracks notes.

The R package allows the user to extract the harmonized data in the long and wide format. The package also provides functions for report generation using the harmonized data.

## Installation

You can install the development version of psHarmonize from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("NUDACC/psHarmonize")
```

## Example

### Harmonization sheet

A harmonization sheet gives instructions to the `harmonization()` function. It can be created in an excel spreadsheet, or within R as a data.frame. It has a row for each combination of variable, cohort, and visit.

An example harmonization sheet is provided with the package.

```{r}
library(psHarmonize)

head(harmonization_sheet_example)

```

### Import source datasets

In order for the harmonization function to harmonize our data, the source datasets must be read into the global environment of your R session.

### Harmonize data

Once your harmonization sheet is created, and your data is loaded, you can call the `harmonization()` function with the harmonization sheet as the input.

(In the following example, the source datasets are already loaded with the R package)

```{r}

harmoniz_obj <- harmonization(harmonization_sheet = harmonization_sheet_example)
  
```

It will create an harmonization object, that will contain a long harmonized dataset (rows for each visit), a wide harmonized dataset (rows for each participant), the input harmonization sheet, and an error log.

The harmonization sheet has an associated `summary()` method which gives some general stats on whether the harmonization was successful.

```{r}

summary(harmoniz_obj)

```

### Extract dataset(s)

The harmonized dataset can then be extracted from the harmonization object.

```{r}

# Extracting long dataset
harmoniz_long_dataset <- harmoniz_obj$long_dataset

# Extracting wide dataset
harmoniz_wide_dataset <- harmoniz_obj$wide_dataset

```

### Create report(s)

There are a few types of reports that can be created from the harmonization object.

#### Error log report

A report from the error log can be created with the `create_error_log_report()` function. This function will create an html file to the path specified.

#### Summary report

A summary report can be created in a similar manner with the `create_summary_report()` function, This function will create descriptive tables and plots of your harmonized data.

#### Summary compare report

If you want to see comparisons between the raw non-harmonized data, and the harmonized data, you can use the argument `compare = TRUE` with the `create_summary_report()` function.

## Reference Paper

Stephen JJ, Carolan P, Krefman AE, et al. psHarmonize: Facilitating reproducible large-scale pre-statistical data harmonization and documentation in R. Patterns (N Y). 2024;5(8):101003. <doi:10.1016/j.patter.2024.101003>
